---
title: "Writeup"
output: html_document
date: "2024-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Chunk one, Yonathan}
#Chunk one, let's use this for basic EDA, initial understanding of the data
housing_data <- read.csv("C:/Users/shime/Downloads/zhvi-all-homes_md-va-dc_2021-2023.csv")
```

# Step 2 - Yonathan

**Break down the dataset by regions.**

```{r Chunk two}
#Using code we group the ZHVIs for the data together by the regions/states, DC, MD, and VA


dc_data <- subset(housing_data, State == "DC")
summary(dc_data)

md_data <- subset(housing_data, State == "MD")
summary(md_data)

va_data <- subset(housing_data, State == "VA")
summary(va_data)
```
Maryland and Virginia are the only of the three that has NA's with a total of 468 NAs'and 1028 NA's. The most expensive home in the dataset is located in Maryland


# Step 3 - Yonathan 

**Finding the average ZHVI of each region over all three years as a whole**
```{r Chunk three} 
#Chunk three, further breakdown 

#Average ZHVI of dc housing over all 3 years
mean_dc <- mean(dc_data$ZHVI, na.rm = TRUE)
mean_dc

#Average ZHVI of va housing over all 3 years
mean_va <- mean(va_data$ZHVI, na.rm = TRUE)
mean_va

#Average ZHVI of md housing over all 3 years
mean_md <- mean(md_data$ZHVI, na.rm = TRUE)
mean_md


```
We can see the average home value from 2021-2023 was the largest in DC (650863), Maryland(420935) and then Virginia(314345)

# Step 4  

**Separating the DC, Maryland, and Virginia data by year**
```{r Chunk 4}
#Create column 'Year' by taking the first 4 characters of the variable 'month'
dc_data$Year <- substr(dc_data$month, 1, 4)
va_data$Year <- substr(va_data$month, 1, 4)
md_data$Year <- substr(md_data$month, 1, 4)

#Creating subsets of dc data by year
dc_data_2021 <- subset(dc_data, Year == "2021")
dc_data_2022 <- subset(dc_data, Year == "2022")
dc_data_2023 <- subset(dc_data, Year == "2023")

#Finding average ZHVI of dc by year
mean_dc_2021 <- mean(dc_data_2021$ZHVI, na.rm = TRUE)
mean_dc_2022 <- mean(dc_data_2022$ZHVI, na.rm = TRUE)
mean_dc_2023 <- mean(dc_data_2023$ZHVI, na.rm = TRUE)

#Creating subsets of VA data by year
va_data_2021 <- subset(va_data, Year == "2021")
va_data_2022 <- subset(va_data, Year == "2022")
va_data_2023 <- subset(va_data, Year == "2023")


#Finding average ZHVI of VA by year
mean_va_2021 <- mean(va_data_2021$ZHVI, na.rm = TRUE)
mean_va_2022 <- mean(va_data_2022$ZHVI, na.rm = TRUE)
mean_va_2023 <- mean(va_data_2023$ZHVI, na.rm = TRUE)


#Creating subsets of MD data by year
md_data_2021 <- subset(md_data, Year == "2021")
md_data_2022 <- subset(md_data, Year == "2022")
md_data_2023 <- subset(md_data, Year == "2023")


#Finding average ZHVI of MD by year
mean_md_2021 <- mean(md_data_2021$ZHVI, na.rm = TRUE)
mean_md_2022 <- mean(md_data_2022$ZHVI, na.rm = TRUE)
mean_md_2023 <- mean(md_data_2023$ZHVI, na.rm = TRUE)
```


# Step 5

**Finding the average ZHVI of each reagion for each given year**
```{r Chunk 5}
#Comparing values from chunk four
options(scipen = 999)

#Average ZHVI of DC for years 2021-2023
mean_dc_2021
mean_dc_2022
mean_dc_2023

#Average ZHVI of VA for years 2021-2023
mean_va_2021
mean_va_2022
mean_va_2023

#Average ZHVI of MD for years 2021-2023
mean_md_2021
mean_md_2022
mean_md_2023
```

From this we observe the average housing price in DC wasn't always increasing it went from 643,242 in 2021 to 662,298 in 2022 and then back down to 647,048 in 2023

Virginia however, was always increasing. Going from 293,828 to 319,719 in 2022 and 328,925 in 2023

Maryland is the same with housing going from 400,198 to 426,621 and 434,221 in 2023.

# Step 6

**Plotting the data from chunk 4 and 5**
```{r Chunk 6}

#Plotting ZHVI for each region by year
library(ggplot2)

#Creating a data frame so all the data is put into one for it to be able to be plotted
combined_data <- rbind(dc_data, md_data, va_data)

ggplot(combined_data, aes(x = substr(month, 1, 4), y = ZHVI, fill = State)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(
    title = "Average Housing Prices (ZHVI) by State/Region Over the Years",
    x = "Year",
    y = "Average ZHVI",
    fill = "State/Region"
  ) +
  scale_y_continuous(
    breaks = seq(200000, 650000, by = 100000)  #Making the scale smaller than the default
  )
```


# Step 7
**First statistical testing on the ZHVI of the three regions from 2021-2023 as a whole**
```{r Chunk 7}
# statistical testing the average home price for all three states/regions using the combined data


anova_result <- aov(ZHVI ~ State, data = combined_data)
summary(anova_result)

tukey_result <- TukeyHSD(anova_result)
tukey_result
```
With a p-value below 0.05, we can reject the null hypothesis that the means of the ZHVI for the three regions/states are equal. After doing a post-hoc test, we can see the average housing price(ZHVI) in Maryland is 229,928 less than DC. Housing price in VA is 336,518 less than DC. And VA is 106,590 cheaper than MD. There are statistically significant differences in ZHVI between all of the regions.

# Step 8

**Statistical testing on the ZHVI of each region for each given year**
```{r Chunk 8}

#Creating sets of data with all regions for each given year within the range of 2021-2023
all_2021 <- rbind(dc_data_2021, va_data_2021, md_data_2021)

all_2022 <- rbind(dc_data_2022, va_data_2022, md_data_2022)

all_2023 <- rbind(dc_data_2023, va_data_2023, md_data_2023)


#Anova test for the year 2021
anova_2021 <- aov(ZHVI ~ State, data = all_2021)
summary(anova_2021)

#Anova test for the year 2022
anova_2022 <- aov(ZHVI ~ State, data = all_2022)
summary(anova_2022)

#Anova test for the year 2023
anova_2023 <- aov(ZHVI ~ State, data = all_2023)
summary(anova_2023)

#Post HOC test for 2021
tukey_result2021 <- TukeyHSD(anova_2021)
tukey_result2021

#Post HOC test for 2022
tukey_result2022 <- TukeyHSD(anova_2022)
tukey_result2022

#Post HOC test for 2023
tukey_result2023 <- TukeyHSD(anova_2023)
tukey_result2023
```
Every year has a p-value below 0.05. 

We observe a 243,045 difference in DC-MD homes, a 349414 difference in VA-DC homes, and a 106369 difference in VA-MD homes for the year 2021. 

For 2022 we observe an avg difference of 235,637 between MD-DC homes, 342579 between VA-DC homes, and 106942 between VA-MD homes.

For 2023 we observe a 212827 diff between MD-DC homes, 318123 for VA-DC homes, and 105296 for VA-MD homes. 

All years and regions had p-values below 0.05 so there are significant differences between all the avg housing prices between every region regardless of the year.

# Step 9 

**Further statistical testing**

```{r Chunk 9}
#ANOVA testing the average ZHVI of a region in one year versus the same region in the other years of the data
#We decided to take this step on Dr.Morris' recommendation following our first presentation of the data. 

anova_dc_all_years <- aov(ZHVI ~ Year, data = dc_data)
summary(anova_dc_all_years)

anova_md_all_years <- aov(ZHVI ~ Year, data = md_data)
summary(anova_md_all_years)

anova_va_all_years <- aov(ZHVI ~ Year, data = va_data)
summary(anova_va_all_years)

```
The p-value for DC is above 0.05, meaning there is no significant difference in the average ZHVI of DC between the years 2021,2022, and 2023. 

The p-value for md and va were both below 0.05. This means we reject the null hypothesis and that there is a difference in the average ZHVI of of MD for at least one of the years (2021-2023) compared to the others. This result is the same for VA. 

```{r}
#Tukey Tests for the p-values below 0.05

tukey_md <- TukeyHSD(anova_md_all_years)
print(tukey_md)
  
tukey_va <- TukeyHSD(anova_va_all_years)
print(tukey_va)

```
The result for maryland confirms that the average ZHVI in the state in the year 2022 was significantly higher than in 2021. The same goes for the average ZHVI in MD from 2023 being significantly higher than 2021. However, the average ZHVI for maryland in 2023 is not significantly higher than the average ZHVI in 2022. 

For Virginia the average ZHVI in 2022 was significantly higher than in 2021. The averge ZHVI in 2023 is significantly higher than in 2021. The average ZHVI in 2023 is also significantly higher than in 2022.

# Step 3- Fardin
** Perform General EDA on the Data set 
```{r}
housing <- read.csv("/Users/fardinhafiz/Desktop/housingdata.csv")
str(housing)
```

```{r}
dc_data <- subset(housing, State == "DC")
md_data <- subset(housing, State == "MD")
va_data <- subset(housing, State == "VA")
```

```{r}
install.packages("ggplot2")
library(ggplot2)
```

```{r}
dc_data$month <- as.Date(dc_data$month)
ggplot(dc_data, aes(x = month, y = ZHVI)) +
    geom_line() +
    labs(title = "Home Value Over Time (DC)", x = "Month", y = "ZHVI (Home Value Index)") +
    scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

```{r}
md_data$month <- as.Date(md_data$month)
ggplot(md_data, aes(x = month, y = ZHVI)) +
    geom_line() +
    labs(title = "Home Value Over Time (MD)", x = "Month", y = "ZHVI (Home Value Index)") +
    scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{R}
va_data$month <- as.Date(va_data$month)
ggplot(va_data, aes(x = month, y = ZHVI)) +
    geom_line() +
    labs(title = "Home Value Over Time (VA)", x = "Month", y = "ZHVI (Home Value Index)") +
    scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

```{r}
ggplot(dc_data, aes(y=ZHVI)) + 
  geom_boxplot( colour="black", fill="blue") +
  labs(title="Boxplot of ZHVI (DC)", x="", y = "ZHVI")+
  theme(plot.title = element_text(face = "bold"),
        axis.title.y = element_text(face = "italic"))
```
```{r}
ggplot(md_data, aes(y=ZHVI)) + 
  geom_boxplot( colour="black", fill="red") +
  labs(title="Boxplot of ZHVI (MD)", x="", y = "ZHVI")+
  theme(plot.title = element_text(face = "bold"),
        axis.title.y = element_text(face = "italic"))
```
```{r}
ggplot(va_data, aes(y=ZHVI)) + 
  geom_boxplot( colour="black", fill="green") +
  labs(title="Boxplot of ZHVI (VA)", x="", y = "ZHVI")+
  theme(plot.title = element_text(face = "bold"),
        axis.title.y = element_text(face = "italic"))

```
# Step 4 Questions- Fardin
What do we know about this dataset?

This dataset is taken from the online housing marketplace Zillow. Zillow is a website that facilitates the sale and purchase
of real estate nationwide. The exploratory question we aimed to answer was if housing prices were increasing in the 
D.C, Maryland, and Virginia area (DMV), and we decided that Zillow would be a very useful tool in studying this 
question. Zillow records all of the purchases and sales made on the website and evaluations of homes on the market
and this provides us an extensive dataset across multiple years that can be evaluated. By analyzing the changes 
in home values across multiple years and multiple areas, it allows us to study our research question. The dataset provided
a lot of valuable information about the areas of interest. The dataset was directly provided by Zillow and is publicly 
available for download. The dataset aims to study the idea of home value by breaking down the evaluation of property into 
zip codes. The dataset takes into account all of the properties and different types of properties in the area and 
provides a ZHVI value (Zillow Home Value Index). This ZHVI value is a measure of the home value of different homes 
in different percentiles of home price. We decided to use the ZHVI that measures between the 35th and 65th percentile
in order to get an understanding of the general average trend of homes. ZHVI can also be specific to different types 
of homes like single-family, co-op, etc but we decided to include all types of homes in our dataset in order to get a full 
understanding of all types of homes in a specific zipcode. This dataset is updated every year but we decided to use the data 
from 2021-2023 in order to get a concept of the recent changes in housing prices. ZHVI is also representative of the actual 
housing sales and is not based on housing trend estimates, so it is more reflective of the actual market. 


What are the limitations of the dataset?
How was the information gathered?
What analysis has already been completed related to the content in your dataset?
How did the research you gathered contribute to your question development?
What additional information would be beneficial?
How did your question change, if at all, after Exploratory Data Analysis?
Based on EDA can you begin to sketch out an answer to your question?
References (APA style preferred)



